---
title: 'Project_Prctice'
author: "Safiya Geelani"
date: "2022-10-07"
output: pdf_document
---

```{r warning = F, message = F}
library(tidyverse)
library(ggplot2)
library(GGally)
library(ggpubr)
library(openintro)
# babies
```

# Read in the data set and basic descriptive statistics
```{r message = F, warning = F}
glimpse(babies)
summary(babies)
```

# Cleaning the data
```{r message = F, warning = F}
babies_data <- babies %>%
  na.omit() %>% 
  mutate_at(vars(smoke, parity), as.factor) %>%
  rename(first_pregnancy = parity) # 0 is first pregnancy
```

# Visualizing the distribution of predictors
```{r message = F, warning = F}
# Histogram for numeric variables
babies_data %>% select_if(is.numeric) %>%
  gather() %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 10) +
  facet_wrap(~ key, scales = "free")

# Barplot for categorical variables
babies_data %>% select_if(negate(is.numeric)) %>%
  gather() %>%
  ggplot(aes(value)) +
  geom_bar() +
  facet_wrap(~ key, scales = "free")

ggpairs(babies_data[,c(8,1:4)])
ggpairs(babies_data[,c(8,5:7)])
```

# Relationship between different variables
```{r message = F, warning = F}
babies_data %>% 
  ggplot() +
  geom_histogram(bins = 20, aes(x = gestation)) + facet_wrap(~smoke)

babies_data %>% 
  ggplot(aes(x = smoke, y = bwt)) +
  geom_boxplot(aes(fill = smoke), show.legend = F)

babies_data %>% 
  ggplot(aes(x = gestation, y = bwt, color = smoke)) +
  geom_point() + geom_smooth(se = F)

babies_data %>% 
  ggplot(aes(x = first_pregnancy, y = age)) +
  geom_boxplot(aes(fill = smoke))
```

# Research questions
1. What are the best predictors of diabetes in this dataset?
2. What is the relationship between probability of diabetes and predictor variables?

# Variable selection
```{r message = F, warning = F}
# Considering full model first
babies_full <- glm(first_pregnancy ~ ., family = "binomial", data = babies_data)
summary(babies_full)

# Final model
babies_red <- update(babies_full, ~. -case, data = babies_data)
summary(babies_red)
```

# Making data frames for doing some predictions
```{r message = F, warning = F}
# Effect of bwt holding other variables constant
babies1 <- with(babies_data, data.frame(bwt = rep(seq(min(bwt), max(bwt),
                                                      length.out = 100),2),
                                      gestation = mean(gestation),
                                      age = mean(age),
                                      height = mean(height),
                                      weight = mean(weight),
                                      smoke = factor(rep(0:1, each = 100))))

# Effect of gestation holding other variables constant
babies2 <- with(babies_data, data.frame(bwt = mean(bwt),
                                      gestation = rep(seq(min(gestation), max(gestation),
                                                          length.out = 100), 2),
                                      age = mean(age),
                                      height = mean(height),
                                      weight = mean(weight),
                                      smoke = factor(rep(0:1, each = 100))))

# Effect of age holding other variables constant
babies3 <- with(babies_data, data.frame(bwt = mean(bwt),
                                      gestation = mean(gestation),
                                      age = rep(seq(min(age), max(age),
                                                    length.out = 100), 2),
                                      height = mean(height),
                                      weight = mean(weight),
                                      smoke = factor(rep(0:1, each = 100))))

# Effect of height holding other variables constant
babies4 <- with(babies_data, data.frame(bwt = mean(bwt),
                                      gestation = mean(gestation),
                                      age = mean(age),
                                      height = rep(seq(min(height), max(height),
                                                       length.out = 100), 2),
                                      weight = mean(weight),
                                      smoke = factor(rep(0:1, each = 100))))

# Effect of weight holding other variables constant
babies5 <- with(babies_data, data.frame(bwt = mean(bwt),
                                      gestation = mean(gestation),
                                      age = mean(age),
                                      height = mean(height),
                                      weight = rep(seq(min(weight), max(weight),
                                                       length.out = 100), 2),
                                      smoke = factor(rep(0:1, each = 100))))
```

# Making predictions
```{r message = F, warning = F}
babies_1 <- cbind(babies1, predict(babies_red, newdata = babies1, type = "link", se = TRUE))
babies_1 <- within(babies_1, {
  pred_prob <- plogis(fit)
  lower <- plogis(fit - (1.96 * se.fit))
  upper <- plogis(fit + (1.96 * se.fit))})

babies_2 <- cbind(babies2, predict(babies_red, newdata = babies2, type = "link", se = TRUE))
babies_2 <- within(babies_2, {
  pred_prob <- plogis(fit)
  lower <- plogis(fit - (1.96 * se.fit))
  upper <- plogis(fit + (1.96 * se.fit))})

babies_3 <- cbind(babies3, predict(babies_red, newdata = babies3, type = "link", se = TRUE))
babies_3 <- within(babies_3, {
  pred_prob <- plogis(fit)
  lower <- plogis(fit - (1.96 * se.fit))
  upper <- plogis(fit + (1.96 * se.fit))})

babies_4 <- cbind(babies4, predict(babies_red, newdata = babies4, type = "link", se = TRUE))
babies_4 <- within(babies_4, {
  pred_prob <- plogis(fit)
  lower <- plogis(fit - (1.96 * se.fit))
  upper <- plogis(fit + (1.96 * se.fit))})

babies_5 <- cbind(babies5, predict(babies_red, newdata = babies5, type = "link", se = TRUE))
babies_5 <- within(babies_5, {
  pred_prob <- plogis(fit)
  lower <- plogis(fit - (1.96 * se.fit))
  upper <- plogis(fit + (1.96 * se.fit))})
```

# Plotting the predictions
```{r message = F, warning = F}
a <- ggplot(babies_1, aes(x = bwt, y = pred_prob)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = smoke), alpha = 0.10) +
  geom_line(aes(colour = smoke), size = 1) + theme_light() +
  xlab("birth_wt") + ylab("Pregnancy_pp") #pp means predicted probabilities

b <- ggplot(babies_2, aes(x = gestation, y = pred_prob)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = smoke), alpha = 0.10) +
  geom_line(aes(colour = smoke), size = 1) +  theme_light() +
  ylab("Pregnancy_pp")

c <- ggplot(babies_3, aes(x = age, y = pred_prob)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = smoke), alpha = 0.10) +
  geom_line(aes(colour = smoke), size = 1) +  theme_light() +
  ylab("Pregnancy_pp")

d <- ggplot(babies_4, aes(x = height, y = pred_prob)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = smoke), alpha = 0.10) +
  geom_line(aes(colour = smoke), size = 1) +  theme_light() +
  ylab("Pregnancy_pp")

e <- ggplot(babies_5, aes(x = weight, y = pred_prob)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = smoke), alpha = 0.10) +
  geom_line(aes(colour = smoke), size = 1) +  theme_light() +
  ylab("Pregnancy_pp")

ggarrange(b, c, d, e, a, 0, ncol = 2, nrow = 3)
```